```{r intro-to-dbi, include = FALSE}
eval_dbi <- FALSE
```

```{r, eval = eval_dbi, include = FALSE}
library(DBI)
library(RSQLite)
```

# Introduction to `DBI` 

## Local database basics
*Connecting and adding data to a database*

1. Load the `DBI` package
    ```{r, eval = TRUE}
    library(DBI)
    ```

1. Use `dbConnect()` to open a Database connection
    ```{r, eval = TRUE}
    con <- dbConnect(RSQLite::SQLite(), "mydatabase.sqlite")
    ```
    
1. Use `dbListTables()` to view existing tables, there should be 0
    ```{r, eval = eval_dbi}
    dbListTables(con)
    ```

1. Use `dbWriteTable()` to create a new table using `mtcars` data
    ```{r, eval = eval_dbi}
    dbWriteTable(con, "mtcars", mtcars)
    ```

1. Use `dbListTables()` to view existing tables, it should return `mtcars`
    ```{r, eval = eval_dbi}
    dbListTables(con)
    ```

1. Use `dbGetQuery()` to pass a SQL query to the database
    ```{r, eval = eval_dbi}
    dbGetQuery(con, "select * from mtcars")
    ```

1. Close the database connection using `dbDisconnect()`
    ```{r, eval = eval_dbi}
    dbDisconnect(con)
    ```

## Options for writing tables
*Understand how certain arguments in `dbWriteTable()` work*

1. Use `dbConnect()` to open a Database connection again
    ```{r, eval = eval_dbi}
    con <- dbConnect(RSQLite::SQLite(), "mydatabase.sqlite")
    ```

1. Use `dbWriteTable()` to create a new table using `mtcars` data
    ```{r, eval = FALSE}
    dbWriteTable(con, "mtcars", mtcars)
    ```
```
## Error: Table mtcars exists in database, and both overwrite and append are FALSE
```

1. Use the `append` argument in `dbWriteTable()` to add the data in the `mtcars` table
    ```{r, eval = eval_dbi}
    dbWriteTable(con, "mtcars", mtcars, append = TRUE)
    ```

1. Using `dbGetQuery()`, check the current record count of `mtcars`, pass the following query: "select count() from mtcars"
    ```{r, eval = eval_dbi}
    dbGetQuery(con, "select count() from mtcars")
    ```

1. Use the `overwrite` argument to `dbWriteTable()` to replace the data in the `mtcars` table
    ```{r, eval = TRUE}
    dbWriteTable(con, "mtcars", mtcars, overwrite = TRUE)
    ```

1. Check the record count of `mtcars` again
    ```{r, eval = eval_dbi}
    dbGetQuery(con, "select count() from mtcars")
    ```

## Database operations
*Understand how to use `dbSendStatement()` to modify the database*

1. Use `dbSendStatement()` to pass a SQL commands that deletes any automatic car form `mtcars`: "delete from mtcars where am = 1". Load the command to a variable called `rs`
    ```{r, eval = eval_dbi}
    rs <- dbSendStatement(con, "delete from mtcars where am = 1")
    ```

1. Call the `rs` variable to view information about the results of the requeted change
    ```{r, eval = eval_dbi}
    rs
    ```

1. Use `dbHasCompleted()` to confirm that the job is complete
    ```{r, eval = eval_dbi}
    dbHasCompleted(rs)
    ```

1. Use `dbGetRowsAffected()` to see the number of rows that were affected by the request
    ```{r, eval = eval_dbi}
    dbGetRowsAffected(rs)
    ```

1. Clear the results using `dbClearResult()`
    ```{r, eval = eval_dbi}
    dbClearResult(rs)
    ```

1. Confirm that the result set has been removed by calling the `rs` variable once more
    ```{r, eval = eval_dbi}
    rs
    ```

1. Check the record count of `mtcars` again, the new count should be 19 (32 original records - 13 deleted records)
    ```{r, eval = eval_dbi}
    dbGetQuery(con, "select count() from mtcars")
    ```

## `knitr` SQL engine
*See how to run SQL queries as code chunks*

1. Start a new code chunk, but using `sql` instead of `r` as the first argument of the chunk.  Also add `connection=con` as another argument of the chunk. 
````markdown
    `r ''````{sql, connection=con}
    select * from mtcars
    ```
````
```{sql, connection=con, echo = FALSE}
select * from mtcars
```

2. Add the `max.print` options to the chunk, and set it to 5
````markdown
    `r ''````{sql, connection=con, max.print = 5}
    select * from mtcars
    ```
````
```{sql, connection=con, echo = FALSE, max.print = 5}
select * from mtcars
```

3. Set defaults for the `sql` chunks by using the `knitr::opts_chunk$set()` command
```{r, eval = TRUE}
knitr::opts_chunk$set(connection = "con", max.print = 5)
```

4. Run the same query in a new `sql` chunk, but without any other argument
````markdown
    `r ''````{sql}
    select * from mtcars
    ```
````
```{sql, echo = FALSE}
select * from mtcars
```

5. Close the database connection using `dbDisconnect()`
```{r, eval = TRUE}
dbDisconnect(con)
```