```{r, intro-to-sparklyr, include = FALSE}
eval_sparklyr <- FALSE
if(Sys.getenv("GLOBAL_EVAL") != "") eval_sparklyr <- Sys.getenv("GLOBAL_EVAL")
```

# Intro to `sparklyr`

```{r, eval = eval_sparklyr, include = FALSE}
library(dplyr)
library(sparklyr)
```

## New Spark session
*Learn to open a new Spark session*

1. Load the `sparklyr` library
    ```{r, eval = eval_sparklyr}
    library(sparklyr)
    ```

1. Use `spark_connect()` to create a new local Spark session
    ```{r, eval = eval_sparklyr}
    sc <- spark_connect(master = "local")
    ```

1. Click on the `Spark` button to view the current Spark session's UI

1. Click on the `Log` button to see the message history

## Data transfer
*Practice uploading data to Spark*

1. Load the `dplyr` library
    ```{r, eval = eval_sparklyr}
    library(dplyr)
    ```

1. Copy the `mtcars` dataset into the session
    ```{r, eval = eval_sparklyr}
    spark_mtcars <- copy_to(sc, mtcars, "my_mtcars")
    ```

1. In the **Connections** pane, expande the `my_mtcars` table

1. Go to the Spark UI, note the new jobs

1. In the UI, click the Storage button, note the new table

1. Click on the **In-memory table my_mtcars** link

## Spark and `dplyr`
*See how Spark handles `dplyr` commands*

1. Run the following code snipett
    ```{r, eval = eval_sparklyr}
    spark_mtcars %>%
      group_by(am) %>%
      summarise(mpg_mean = mean(mpg, na.rm = TRUE))
    ```

1. Go to the Spark UI and click the **SQL** button 

1. Click on the top item inside the **Completed Queries** table

1. At the bottom of the diagram, expand **Details**



