```{r intro-to-dbi, include = FALSE}
eval_dbi <- FALSE
if(Sys.getenv("GLOBAL_EVAL") != "") eval_dbi <- Sys.getenv("GLOBAL_EVAL")
```

```{r, eval = TRUE, include = FALSE}
library(DBI)
```

# Introduction to `DBI` 

## Local database basics
*Connecting and adding data to a database*

1. Load the `DBI` package
    ```{r, eval = TRUE}
    library(DBI)
    ```

1. Use `dbConnect()` to open a Database connection
    ```{r, eval = TRUE}
    con <- dbConnect(RSQLite::SQLite(), "mydatabase.sqlite")
    ```
    
1. Use `dbListTables()` to view existing tables, there should be 0 tables
    ```{r, eval = eval_dbi}
    dbListTables(con)
    ```

1. Use `dbWriteTable()` to create a new table using `mtcars` data. Name it **db_mtcars**
    ```{r, eval = eval_dbi}
    dbWriteTable(con, "db_mtcars", mtcars)
    ```

1. Use `dbListTables()` to view existing tables, it should return **db_mtcars**
    ```{r, eval = eval_dbi}
    dbListTables(con)
    ```

1. Use `dbGetQuery()` to pass a SQL query to the database
    ```{r, eval = eval_dbi}
    dbGetQuery(con, "select * from db_mtcars")
    ```

1. Close the database connection using `dbDisconnect()`
    ```{r, eval = eval_dbi}
    dbDisconnect(con)
    ```

## Options for writing tables
*Understand how certain arguments in `dbWriteTable()` work*

1. Use `dbConnect()` to open a Database connection again
    ```{r, eval = eval_dbi}
    con <- dbConnect(RSQLite::SQLite(), "mydatabase.sqlite")
    ```

1. Use `dbWriteTable()` to re-create the **db_mtcars** table using `mtcars` data
    ```{r, eval = FALSE}
    dbWriteTable(con, "db_mtcars", mtcars)
    ```
```
## Error: Table db_mtcars exists in database, and both overwrite and append are FALSE
```

1. Use the `append` argument in `dbWriteTable()` to add the data in the **db_mtcars** table
    ```{r, eval = eval_dbi}
    dbWriteTable(con, "db_mtcars", mtcars, append = TRUE)
    ```

1. Using `dbGetQuery()`, check the current record count of **db_mtcars**, pass the following query: "select count() from db_mtcars"
    ```{r, eval = eval_dbi}
    dbGetQuery(con, "select count() from db_mtcars")
    ```

1. Use the `overwrite` argument to `dbWriteTable()` to replace the data in the **db_mtcars** table
    ```{r, eval = TRUE}
    dbWriteTable(con, "db_mtcars", mtcars, overwrite = TRUE)
    ```

1. Check the record count of `db_mtcars` again
    ```{r, eval = eval_dbi}
    dbGetQuery(con, "select count() from db_mtcars")
    ```

## Database operations
*Understand how to use `dbSendStatement()` to modify the database*

1. Use `dbSendStatement()` to pass a SQL commands that deletes any automatic car form `db_mtcars`: "delete from db_mtcars where am = 1". Load the command to a variable called `rs`
    ```{r, eval = eval_dbi}
    rs <- dbSendStatement(con, "delete from db_mtcars where am = 1")
    ```

1. Call the `rs` variable to view information about the results of the requested change
    ```{r, eval = eval_dbi}
    rs
    ```

1. Use `dbHasCompleted()` to confirm that the job is complete
    ```{r, eval = eval_dbi}
    dbHasCompleted(rs)
    ```

1. Use `dbGetRowsAffected()` to see the number of rows that were affected by the request
    ```{r, eval = eval_dbi}
    dbGetRowsAffected(rs)
    ```

1. Clear the results using `dbClearResult()`
    ```{r, eval = eval_dbi}
    dbClearResult(rs)
    ```

1. Confirm that the result set has been removed by calling the `rs` variable once more
    ```{r, eval = eval_dbi}
    rs
    ```

1. Check the record count of `db_mtcars` again, the new count should be 19 (32 original records - 13 deleted records)
    ```{r, eval = eval_dbi}
    dbGetQuery(con, "select count() from db_mtcars")
    ```

## `knitr` SQL engine
*See how to run SQL queries as code chunks*

1. Start a new code chunk, but using `sql` instead of `r` as the first argument of the chunk.  Also add `connection=con` as another argument of the chunk. 
````markdown
    `r ''````{sql, connection=con}
    select * from db_mtcars
    ```
````
```{sql, connection=con, echo = FALSE}
select * from db_mtcars
```

1. Add the `max.print` options to the chunk, and set it to 5
````markdown
    `r ''````{sql, connection=con, max.print = 5}
    select * from db_mtcars
    ```
````
```{sql, connection=con, echo = FALSE, max.print = 5}
select * from db_mtcars
```

1. Set defaults for the `sql` chunks by using the `knitr::opts_chunk$set()` command
```{r, eval = TRUE}
knitr::opts_chunk$set(connection = "con", max.print = 5)
```

1. Run the same query in a new `sql` chunk, but without any other argument
````markdown
    `r ''````{sql}
    select * from db_mtcars
    ```
````
```{sql, echo = FALSE}
select * from db_mtcars
```

1. Store the results of the query into an R object called **local_mtcars** using 
the `output.var` option.
```{sql, output.var = "local_mtcars"}
select * from db_mtcars
```

```{r}
local_mtcars
```

1. Close the database connection using `dbDisconnect()`
```{r, eval = TRUE}
dbDisconnect(con)
```
