```{r db-advanced, include = FALSE}
eval_adv <- FALSE
if(Sys.getenv("GLOBAL_EVAL") != "") eval_adv <- Sys.getenv("GLOBAL_EVAL")
```

```{r, eval = eval_adv, include = FALSE}
library(connections)
library(dplyr)
library(rlang)
library(config)
library(dbplyr)
library(purrr)
```

# Advanced Operations

## Simple wrapper function

1. Load the `connections` and `dplyr` libraries
    ```{r, eval = eval_adv}
    library(connections)
    library(dplyr)
    library(dbplyr)
    library(config)
    ```

1. Use `connection_open()` to open a Database connection
    ```{r, eval = eval_adv}
    con <- connection_open(
      RPostgres::Postgres(),
      host =  "localhost",
      user = get("user"),
      password = get("pwd"),
      port = 5432,
      dbname = "postgres",
      bigint = "integer"
    )
    ```

1. Create a variable that points to the **v_orders** table
    ```{r, eval = eval_adv}
    orders <- tbl(con, in_schema("retail", "v_orders"))
    ```


1. Create a simple `dplyr` call that gets the average of all order totals
    ```{r, eval = eval_adv}
    orders %>%
      summarise(mean = mean(order_total, na.rm = TRUE))
    ```

1. Load the `rlang` library
    ```{r, eval = eval_adv}
    library(rlang)
    ```

1. Create a new function call `my_mean()` that will take an argument, `x`, and then returns the results of `enquo(x)`
    ```{r, eval = eval_adv}
    my_mean <- function(x){
      enquo(x)
    }
    ```

1. Test the new function. It should return the same variable name, but inside quosure. Use `order_total` as its argument's value to test
    ```{r, eval = eval_adv}
    my_mean(order_total)
    ```

1. In the function, re-assign `x` to the result of `enquo(x)`, and then return `x`
    ```{r, eval = eval_adv}
    my_mean <- function(x){
      x <- enquo(x)
      x
    }
    ```

1. Test the same way again, the output should match to what it was as before
    ```{r, eval = eval_adv}
    my_mean(order_total)
    ```

1. Remove the last line that has `x`,  add the contents of the function with the initial `dplyr` code from step 3.  Then replace `order_total` with `!! x`
    ```{r, eval = eval_adv}
    my_mean <- function(x){
      x <- enquo(x)
      orders %>%
        summarise(mean = mean(!! x, na.rm = TRUE))
    }
    ```

1. Test the new function by passing `order_total` as `x`
    ```{r, eval = eval_adv}
    my_mean(order_total)
    ```

1. In the `summarise()` step, replace the name `mean`, with `!! as_label(x)`, also replace the `=` sign, with `:=`
    ```{r, eval = eval_adv}
    my_mean <- function(x){
      x <- enquo(x)
      orders %>%
        summarise(!! as_label(x) := mean(!! x, na.rm = TRUE))
    }
    ```

1. Run the function again, the name of the column should match the argument value
    ```{r, eval = eval_adv}
    my_mean(order_total)
    ```

1. Test the function by passing a formula, such as `order_total / order_qty`

    ```{r, eval = eval_adv}
    my_mean(order_total / order_qty)
    ```

1. Make the function generic, add a new argument called: `.data`. Inisde the function, replace `orders` with `.data`
    ```{r, eval = eval_adv}
    my_mean <- function(.data, x){
      x <- enquo(x)
      .data %>%
        summarise(!! as_label(x) := mean(!! x, na.rm = TRUE))
    }
    ```

1. The function now behaves more like a `dplyr` verb. Start with `orders` and then pipe into the function
    ```{r, eval = eval_adv}
    orders %>% 
      my_mean(order_total)
    ```

1. Clean up the code by removing the pipe that inside the function
    ```{r, eval = eval_adv}
    my_mean <- function(.data, x){
      x <- enquo(x)
      summarise(
        .data,
        !! as_label(x) := mean(!! x, na.rm = TRUE)
        )
    }
    ```

1. Confirm that there is no change in the behavior of the function
    ```{r, eval = eval_adv}
    orders %>% 
      my_mean(order_total)
    ```

1. Add a `show_query()` step to preview the resulting SQL statement
    ```{r, eval = eval_adv}
    orders %>% 
      my_mean(order_total) %>%
      show_query()
    ```

1. Try the function with a non-DB backed variable, such as `mtcars`. Use `mpg` as the aggregating variable
    ```{r, eval = eval_adv}
    mtcars %>% 
      my_mean(mpg)
    ```

## Multiple variables
*Create functions that handle a variable number of arguments. The goal of the exercise is to create an `anti-select()` function.*

1. Load the `purrr` package
    ```{r, eval = eval_adv}
    library(purrr)
    ```


1. Use *...* as the second argument of a function called `de_select()`.  Inside the function use `enquos()` to parse it
    ```{r, eval = eval_adv}
    de_select <- function(.data, ...){
      vars <- enquos(...)
      vars
    }
    ```

1. Test the function using *airports*
    ```{r, eval = eval_adv}
    orders %>%
      de_select(order_id, date)
    ```

1. Add a step to the function that iterates through each quosure and prefixes a minus sign to tell `select()` to drop that specific field.  Use `map()` for the iteration, and `quo()` to create the prefixed expression.
    ```{r, eval = eval_adv}
    de_select <- function(.data, ...){
      vars <- enquos(...)
      vars <- map(vars, ~ quo(- !! .x))
      vars
    }
    ```

1. Run the same test to view the new results
    ```{r, eval = eval_adv}
    orders %>%
      de_select(order_id, date)
    ```

1. Add the `select()` step.  Use *!!!* to parse the *vars* variable inside `select()`
    ```{r, eval = eval_adv}
    de_select <- function(.data, ...){
      vars <- enquos(...)
      vars <- map(vars, ~ quo(- !! .x))
      select(.data, !!! vars)
    }
    ```

1. Run the test again, this time the operation will take place.  
    ```{r, eval = eval_adv}
    orders %>%
      de_select(order_id, date)
    ```

1. Add a `show_query()` step to see the resulting SQL
    ```{r, eval = eval_adv}
    orders %>%
      de_select(order_id, date) %>%
      show_query()
    ```

1. Test the function with a different data set, such as `mtcars`
    ```{r, eval = eval_adv}
    mtcars %>%
      de_select(mpg, wt, am)
    ```

## Multiple queries
*Suggested approach to avoid passing multiple, and similar, queries to the database*

1. Create a simple `dplyr` piped operation that returns the mean of *order_total* for the months of January, February and March as a group
    ```{r, eval = eval_adv}
    orders %>%
      filter(date_month %in% c(1,2,3)) %>%
      summarise(mean = mean(order_total, na.rm = TRUE)) 
    ```

1. Assign the first operation to a variable called *a*, and create copy of the operation but changing the selected months to January, March and April.  Assign the second one to a variable called *b*.
    ```{r, eval = eval_adv}
    a <- orders %>%
      filter(date_month %in% c(1,2,3)) %>%
      summarise(mean = mean(order_total, na.rm = TRUE)) 
    
    b <- orders %>%
      filter(date_month %in% c(1,3,4)) %>%
      summarise(mean = mean(order_total, na.rm = TRUE)) 
    ```

1. Use *union()* to pass *a* and *b* at the same time to the database
    ```{r, eval = eval_adv}
    union(a, b)
    ```

1. Pipe the previous instruction to `show_query()` to confirm that the resulting query is a single one
    ```{r, eval = eval_adv}
    union(a, b) %>%
      show_query()
    ```


1. Assign to a new variable called *months* an overlapping set of months
    ```{r, eval = eval_adv}
    months <- list(
      c(1,2,3),
      c(1,3,4),
      c(2,4,6)
    )
    ```

1. Use `map()` to cycle through each set of overlapping months.  Notice that it returns three separate results, meaning that it went to the database three times
    ```{r, eval = eval_adv}
    months %>%
      map(
        ~ orders %>%
            filter(date_month %in% .x) %>%
            summarise(mean = mean(order_total, na.rm = TRUE)) 
      )
    ```

1. Add a `reduce()` operation and use `union()` command to create a single query
    ```{r, eval = eval_adv}
    months %>%
      map(
        ~ orders %>%
            filter(date_month %in% .x) %>%
            summarise(mean = mean(order_total, na.rm = TRUE)) 
      ) %>%
      reduce(function(x, y) union(x, y))
    ```

1. Use `show_query()` to see the resulting single query sent to the database
    ```{r, eval = eval_adv}
    months %>%
      map(
        ~ orders %>%
            filter(date_month %in% .x) %>%
            summarise(mean = mean(order_total, na.rm = TRUE)) 
      ) %>%
      reduce(function(x, y) union(x, y)) %>%
      show_query()
    ```


## Multiple queries with an overlapping range

1. Create a table with a *from* and *to* ranges
    ```{r, eval = eval_adv}
    ranges <- tribble(
      ~ from, ~to, 
           1,   4,
           2,   5,
           3,   7
    )
    ```

1. See how `map2()` works by passing the two variables as the *x* and *y* arguments, and adding them as the function
    ```{r, eval = eval_adv}
    map2(ranges$from, ranges$to, ~.x + .y)
    ```

1. Replace *x + y* with the `dplyr` operation from the previous exercise.  In it, re-write the filter to use *x* and *y* as the month ranges 
    ```{r, eval = eval_adv}
    map2(
      ranges$from, 
      ranges$to,
       ~ orders %>%
            filter(date_month >= .x & date_month <= .y) %>%
            summarise(mean = mean(order_total, na.rm = TRUE)) 
    )
    ```

1. Add the `reduce()` operation
    ```{r, eval = eval_adv}
    map2(
      ranges$from, 
      ranges$to,
       ~ orders %>%
            filter(date_month >= .x & date_month <= .y) %>%
            summarise(mean = mean(order_total, na.rm = TRUE)) 
    ) %>%
    reduce(function(x, y) union(x, y))
    ```

5. Add a `show_query()` step to see how the final query was constructed.
    ```{r, eval = eval_adv}
    map2(
      ranges$from, 
      ranges$to,
       ~ orders %>%
            filter(date_month >= .x & date_month <= .y) %>%
            summarise(mean = mean(order_total, na.rm = TRUE)) 
    ) %>%
    reduce(function(x, y) union(x, y)) %>%
    show_query()
    ```


## Characters to field names

1. Create two character variables. One with the name of a field in *flights* and another with a new name to be given to the field
    ```{r, eval = eval_adv}
    my_field <- "new"
    orders_field <- "order_total"
    ```

1. Add a `mutate()` step that adds the new field. And then another step selecting just the new field
    ```{r, eval = eval_adv}
    orders %>%
      mutate(my_field = !! orders_field) %>%
      select(my_field)
    ```

1. Add a `mutate()` step that adds the new field. And then another step selecting just the new field
    ```{r, eval = eval_adv}
    orders %>%
      mutate(!! my_field := !! orders_field) %>%
      select(my_field)
    ```

1. Wrap `orders_field` inside a `sym()` function
    ```{r, eval = eval_adv}
    orders %>%
      mutate(!! my_field := !! sym(orders_field)) %>%
      select(my_field)
    ```

1. Pipe the code into `show_query()`
    ```{r, eval = eval_adv}
    orders %>%
      mutate(!! my_field := !! sym(orders_field)) %>%
      select(my_field) %>%
      show_query()
    ```

```{r, eval = eval_adv, include = FALSE}
connection_close(con)
```
