```{r db-advanced, include = FALSE}
eval_adv <- TRUE
if(Sys.getenv("GLOBAL_EVAL") != "") eval_adv <- Sys.getenv("GLOBAL_EVAL")
```

```{r, eval = eval_adv, include = FALSE}
library(connections)
library(dplyr)
library(rlang)
library(config)
library(dbplyr)
```

# Advanced Operations

## Simple wrapper function

1. Load the `connections` and `dplyr` libraries
    ```{r, eval = eval_adv}
    library(connections)
    library(dplyr)
    library(dbplyr)
    library(config)
    ```

1. Use `connection_open()` to open a Database connection
    ```{r, eval = eval_adv}
    con <- connection_open(
      RPostgres::Postgres(),
      host =  "localhost",
      user = get("user", config = "datawarehouse"),
      password = get("password", config = "datawarehouse"),
      port = 5432,
      dbname = "postgres",
      bigint = "integer"
    )
    ```

1. Create a variable that points to the **v_orders** table
    ```{r, eval = eval_adv}
    orders <- tbl(con, in_schema("datawarehouse", "v_orders"))
    ```


1. Create a simple `dplyr` call that gets the average of all order totals
    ```{r, eval = eval_adv}
    orders %>%
      summarise(mean = mean(order_total, na.rm = TRUE))
    ```

1. Load the `rlang` library
    ```{r, eval = eval_adv}
    library(rlang)
    ```

1. Create a new function call `my_mean()` that will take an argument, `x`, and then returns the results of `enquo(x)`
    ```{r, eval = eval_adv}
    my_mean <- function(x){
      enquo(x)
    }
    ```

1. Test the new function. It should return the same variable name, but inside quosure. Use `order_total` as its argument's value to test
    ```{r, eval = eval_adv}
    my_mean(order_total)
    ```

1. In the function, re-assign `x` to the result of `enquo(x)`, and then return `x`
    ```{r, eval = eval_adv}
    my_mean <- function(x){
      x <- enquo(x)
      x
    }
    ```

1. Test the same way again, the output should match to what it was as before
    ```{r, eval = eval_adv}
    my_mean(order_total)
    ```

1. Remove the last line that has `x`,  add the contents of the function with the initial `dplyr` code from step 3.  Then replace `order_total` with `!! x`
    ```{r, eval = eval_adv}
    my_mean <- function(x){
      x <- enquo(x)
      orders %>%
        summarise(mean = mean(!! x, na.rm = TRUE))
    }
    ```

1. Test the new function by passing `order_total` as `x`
    ```{r, eval = TRUE}
    my_mean(order_total)
    ```

1. In the `summarise()` step, replace the name `mean`, with `!! as_label(x)`, also replace the `=` sign, with `:=`
    ```{r, eval = eval_adv}
    my_mean <- function(x){
      x <- enquo(x)
      orders %>%
        summarise(!! as_label(x) := mean(!! x, na.rm = TRUE))
    }
    ```

1. Run the function again, the name of the column should match the argument value
    ```{r, eval = eval_adv}
    my_mean(order_total)
    ```

1. Test the function by passing a formula, such as `order_total / order_qty`

    ```{r, eval = eval_adv}
    my_mean(order_total / order_qty)
    ```

1. Make the function generic, add a new argument called: `.data`. Inisde the function, replace `orders` with `.data`
    ```{r, eval = eval_adv}
    my_mean <- function(.data, x){
      x <- enquo(x)
      .data %>%
        summarise(!! as_label(x) := mean(!! x, na.rm = TRUE))
    }
    ```

1. The function now behaves more like a `dplyr` verb. Start with `orders` and then pipe into the function
    ```{r, eval = eval_adv}
    orders %>% 
      my_mean(order_total)
    ```

1. Clean up the code by removing the pipe that inside the function
    ```{r, eval = eval_adv}
    my_mean <- function(.data, x){
      x <- enquo(x)
      summarise(
        .data,
        !! as_label(x) := mean(!! x, na.rm = TRUE)
        )
    }
    ```

1. Confirm that there is no change in the behavior of the function
    ```{r, eval = eval_adv}
    orders %>% 
      my_mean(order_total)
    ```

1. Add a `show_query()` step to preview the resulting SQL statement
    ```{r, eval = eval_adv}
    orders %>% 
      my_mean(order_total) %>%
      show_query()
    ```

1. Try the function with a non-DB backed variable, such as `mtcars`. Use `mpg` as the aggregating variable
    ```{r, eval = eval_adv}
    mtcars %>% 
      my_mean(mpg)
    ```

## Multiple variables
## Multiple queries
## Characters to field names
